---
title: "Workplace trends and insight on Fritolay"
author: "Laura Ahumada and Alan Abadzic"
date: "10/5/2021"
output: html_document
---
  



```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


```{r include=FALSE}
#Load Libraries
library(scales)  
library(gridExtra)
library(dplyr)
library(tidyr)
library(stringi)
library(rvest) #html_table, html_node
library(ggplot2)
library(RCurl) #getURL
library(naniar)
library(e1071)
library(tm) 
library(stringr)
library("kableExtra")
library(ggthemes)
library("ggpubr")
library(class)
library(caret)
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer for visualization
library(ggstatsplot)
library(e1071)
library(caret)
```


- All attributes in the data
```{r}
#Reading Files
fritolay=read.csv("CaseStudy2-data.csv")
# all columns
names(fritolay)

```


# Exploratory Analysis on the data
## EDA: Categorical values
```{r}
#Missing values
missing=sapply(fritolay, function(x) sum(is.na(x))) #no missing values
#data with descriptive values
char_data=fritolay[, sapply(fritolay, class) == 'character']  #9 char values
#data with continuous values
cont_data=fritolay[, sapply(fritolay, class) != 'character']  #27 num values

# unique values in each descriptive value
unique_col=char_data %>% summarize_all(n_distinct)
unique_col=gather(unique_col, Columns_names, Unique_Values, Attrition:OverTime) %>% arrange(Unique_Values)
unique_col
# Only JobRole and EducationField have more than 3 unique characters

# unique var for all char variable
unique_chars=sapply(char_data, function(x) (unique(x))) #no empty strings or or anything
#remove Over18 there is only one value. Not useful
fritolay=fritolay %>% select(-Over18)
```

- No NA's nor empty values in Categorical data 
- Only Educational Field and Job Role have more than 3 unique attribute
- All people are over 18 years old
 - Over18 columns can be removed as it has a yes for the whole data

\center __Percent distribution__ \center

```{r, echo=FALSE, out.width="180%", out.height="150%"}

f_m=fritolay %>% group_by(Gender) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot1=f_m %>% ggplot(aes(x=Gender, y=percent, fill=Gender)) + 
  ylim(0,100)+
  geom_col()+ggtitle("Gender")   +
  geom_text(data=f_m, size=3, aes(x=Gender, label=(paste((as.integer(percent)),"%",sep="")), y=percent, fill=NULL), nudge_y=8)+
  coord_polar() +
  theme(axis.text = element_text(size = 8)) + theme_void() 

M_s=fritolay %>% group_by(MaritalStatus) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot2=M_s %>% ggplot(aes(x=MaritalStatus, y=percent, fill=MaritalStatus)) + 
  ylim(0,100)+
  geom_col()+ggtitle("Marital Status")  + theme(legend.position = "none") +
  geom_text(data=M_s, size=3, aes(x=MaritalStatus, label=(paste((as.integer(percent)),"%",sep="")), y=percent, fill=NULL), nudge_y=8)+
  coord_polar() + theme_void() +
  theme(legend.title = element_blank()) +theme(plot.title = element_text(hjust = 0.5))

attr=fritolay %>% group_by(Attrition) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot3=attr %>% ggplot(aes(x=Attrition, y=percent, fill=Attrition)) + 
  ylim(0,100)+
  geom_col()+ggtitle("Attrition")  +
  geom_text(data=attr, size=3, aes(x=Attrition, label=(paste((as.integer(percent)),"%",sep="")), y=percent, fill=NULL), nudge_y=8)+
  coord_polar() +
  theme(axis.text = element_text(size = 8)) + theme_void() +theme(plot.title = element_text(hjust = 0.5)) 

ot=fritolay %>% group_by(OverTime) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot4=ot %>% ggplot(aes(x=OverTime, y=percent, fill=OverTime)) + 
  ylim(0,100)+
  geom_col()+ggtitle("OverTime")  + theme(legend.position = "none") +
  geom_text(data=ot, size=3, aes(x=OverTime, label=(paste((as.integer(percent)),"%",sep="")), y=percent, fill=NULL), nudge_y=6)+
  coord_polar() +
  theme(axis.text = element_text(size = 8)) + theme_void() +
  theme(legend.title = element_blank()) 

bt=fritolay %>% group_by(BusinessTravel) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot5=bt %>% ggplot(aes(x=BusinessTravel, y=percent, fill=BusinessTravel)) +
  ylim(0,100)+
  geom_col()+ggtitle("Travel")  + theme(legend.position = "none") +
  geom_text(data=bt,size=3, aes(x=BusinessTravel,label=(paste((as.integer(percent)),"%",sep="")), y=percent, fill=NULL), nudge_y=6)+
  coord_polar()+
  theme(axis.text = element_text(size = 6)) + theme_void()+
  theme(legend.title = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5))

dp=fritolay %>% group_by(Department) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot6=dp %>% ggplot(aes(x=Department, y=percent, fill=Department)) + 
  ylim(0,100)+
  geom_col()+ggtitle("Department")   +
  geom_text(data=dp,size=3, aes(x=Department, label=(paste((as.integer(percent)),"%",sep="")), fill=NULL), nudge_y=8)+
  coord_polar() +
  theme(axis.text = element_text(size = 6)) + theme_void() +
 theme(plot.title = element_text(hjust = 0.5))  +
  theme(legend.title = element_blank()) 

dp=fritolay %>% group_by(JobInvolvement) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot7=dp %>% ggplot(aes(x=as.factor(JobInvolvement), y=percent, fill=as.factor(JobInvolvement))) + 
  ylim(0,100)+
  geom_col()+ggtitle("Job Involvement")   +
  geom_text(data=dp,size=3, aes(x=as.factor(JobInvolvement), label=(paste((as.integer(percent)),"%",sep="")), fill=NULL), nudge_y=8)+
  coord_polar() +
  theme(axis.text = element_text(size = 6)) + theme_void() +
 theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.title = element_blank()) 

dp=fritolay %>% group_by(JobSatisfaction) %>%
  dplyr::summarize(count=n()) %>% mutate(percent = (count / sum(count))*100)
plot8=dp %>% ggplot(aes(x=as.factor(JobSatisfaction), y=percent, fill=as.factor(JobSatisfaction))) + 
  ylim(0,100)+
  geom_col()+ggtitle("Job Satisfaction")  + 
  geom_text(data=dp,size=3, aes(x=as.factor(JobSatisfaction), label=(paste((as.integer(percent)),"%",sep="")), fill=NULL), nudge_y=8)+
  coord_polar() +
  theme(axis.text = element_text(size = 6)) + theme_void()+
  theme(legend.title = element_blank()) +
  theme(plot.title = element_text(hjust = 0.5))


#job involvement-job satisfaction

grid.arrange(plot3, plot1, plot4,plot2, plot5,plot6, ncol=2)
```

```{r, echo=FALSE, fig.align = 'center'}
grid.arrange(plot7, plot8, ncol=2)

```
\center __Features vs attrition percent__ \center


```{r, echo=FALSE, fig.align = 'center'}
m=fritolay %>% group_by(MaritalStatus) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)

a=m %>% ggplot(aes(x =MaritalStatus, y=per, fill=Attrition)) +
  geom_col() +ggtitle("MaritalStatus") +
  labs(y = "Count")+xlab("MaritalStatus")+
  theme_classic() +
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), y = per + 0.15), vjust = -.3) +
  theme(legend.position="none")

#-----
m=fritolay %>% group_by(BusinessTravel) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)

b=m %>% ggplot(aes(x =per, y=BusinessTravel, fill=Attrition)) + 
  geom_col() +xlab("Percent")+
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), x = per), position = position_stack(vjust = 0.4)) +
  theme_classic() +
  ylab("Travel")+ggtitle("Business Travel")+ theme(plot.title = element_text(hjust = 0.5)) 

#b=fritolay %>% ggplot(aes(y = as.factor(BusinessTravel))) +
 # geom_bar(aes(fill = Attrition)) +ggtitle("BusinessTravel") +
  #ylab("BusinessTravel") +labs(x = "Count")+
#  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), x = per + 0.15), nudge_x=2, vjust = -.3)+theme(legend.position = "none")
#---------
m=fritolay %>% group_by(Department) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)

c=m %>% ggplot(aes(x =per, y=Department, fill=Attrition)) + 
  geom_col() +xlab("Percent")+
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), x = per), position = position_stack(vjust = 0.4)) +
  theme_classic() +
  ylab("Department")+theme(legend.position = "none")

#--------
m=fritolay %>% group_by(EducationField) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)

d=m %>% ggplot(aes(x =per, y=EducationField, fill=Attrition)) + 
  geom_col() +xlab("Percent")+
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), x = per), position = position_stack(vjust = 0.4)) +
  theme_classic() +
  ylab("EducationField")+theme(legend.position = "none")+ theme(axis.text = element_text(size = 6)) 

#---------
m=fritolay %>% group_by(Gender) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)

e=m %>% ggplot(aes(x =Gender, y=per, fill=Attrition)) +
  geom_col() +ggtitle("Gender") +
  labs(y = "Count")+xlab("Gender")+
  theme_classic() +
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), y = per + 0.15), vjust = -.3) +
  theme(legend.position="none")

#---------
m=fritolay %>% group_by(JobRole) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)


f=m %>% ggplot(aes(x =per, y=JobRole, fill=Attrition)) + 
  geom_col() +xlab("Percent")+
  geom_text(size=2,data=m, aes(label=paste(per,"%", sep=""), x = per), position = position_stack(vjust = 0.4)) +
  theme_classic() +
  ylab("Job Role")+theme(legend.position = "none")+ theme(axis.text = element_text(size = 5)) 

#f=fritolay %>% ggplot(aes(y = as.factor(JobRole))) +
  ##geom_bar(aes(fill = Attrition)) +ggtitle("JobRole") +
  #labs(y = "Count")+
  #geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), x = per), position = position_stack(vjust = 0.4)) +
  #ylab("JobRole")+theme(legend.position = "none")
#---------
m=fritolay %>% group_by(OverTime) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=100-(round(((sum(Attrition=="No")/n())*100),1)))
m=gather(m, Attrition, per,Yes:No, factor_key=TRUE)


g=m %>% ggplot(aes(x =OverTime, y=per, fill=Attrition)) +
  geom_col() +ggtitle("Over Time") +
  labs(y = "Count")+xlab("Over Time")+
  theme_classic() +
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), y = per + 0.15), vjust = -.3) 

#--------

#job Satisfaction
j=fritolay %>% group_by(JobSatisfaction) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=round((100-((sum(Attrition=="No")/n())*100)),1))
m=gather(j, Attrition, per,Yes:No, factor_key=TRUE)

js=m %>% ggplot(aes(x =as.factor(JobSatisfaction), y=per, fill=Attrition)) +
  geom_col() +ggtitle("Job Satisfaction")+xlab("Job Satisfaction")+
  theme_classic() +
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), y = per + 0.15), vjust = -.3) 

j=fritolay %>% group_by(JobInvolvement) %>% summarize(No=round(((sum(Attrition=="No")/n())*100),1),Yes=round((100-((sum(Attrition=="No")/n())*100)),1))
m=gather(j, Attrition, per,Yes:No, factor_key=TRUE)

#ji=fritolay %>% ggplot(aes(x = as.factor(JobInvolvement))) +
 # geom_bar(aes(fill = Attrition,position = "dodge")) +ggtitle("Attrition by Job Involvement") +
  #xlab("Job Involvement") +labs(y = "Count")+
  #geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), y = per+.15), position = position_stack(vjust = 0.9))  

ji=m %>% ggplot(aes(x =as.factor(JobInvolvement), y=per, fill=Attrition)) +
  geom_col() +ggtitle("Job Involvement") +
  labs(y = "Count")+xlab("Job Involvement")+
  theme_classic() +
  geom_text(size=3,data=m, aes(label=paste(per,"%", sep=""), y = per + 0.15), vjust = -.3) + theme(legend.position="none")

grid.arrange(a,e,g, ncol=3)
grid.arrange(ji, js, ncol=2)

```
#---------------------------------------------------------------------

```{r, echo=FALSE, out.width="100%", out.height="150%",  fig.align = 'center'}
grid.arrange(b,c,f,d, ncol=1)
```
**Insights for categorical variables**

- MaritalStatus vs churn
  - 47% of the people are married, followed by single 30% and then divorce 21%
  - Single people have the highest overturn rate of 26% and divorce has the lowest with 6%
## People that do over time have the highest highest overturn rate of 68% more churn than those that don't do over time
  - Those that do Overtime have 32% churn while those that don't have 9.7%.
  - Human Resources have the lowest amount of people but the highest churn percent, 27%
  - HR only has 4%, that is very little. Maybe a survey should also be performed to see if people more HR support is needed to make employees feel confortable.
  
  

## EDA: Continuos Variables

```{r, echo=FALSE}

# Percent of zero's in each of the ontinues/numeric columns 
contin_zero=sapply(fritolay, function(x) (((sum(x==0))/870)*100)) 
# make into a data frame
contin_zero=as.data.frame(contin_zero)
contin_zero$Variable=rownames(contin_zero)
rownames(contin_zero)<-c()
#plot to visualize
contin_zero %>% ggplot(aes(x=contin_zero, y=Variable, fill=Variable))+ 
  geom_col() +theme(legend.position = "none") +
  xlab("Percent of zero's") +
  ggtitle("Percent of zero's in data") + xlim(0,100)+
  geom_text(size=3,position = position_dodge(width= .5), aes(label=paste((as.integer(contin_zero)),"%", sep="")),vjust= .5, hjust = -0.5) 

```

- Almost 50% of the data with stock option a zero tells that many people opt out or are not given the option of Stock-option.
- Per the amount of zero's in promotion, it is showing that within a year many people have been promoted.
- Per the zero years with current manager and year in role, we can see that there are quite a few new workers


- Let's see distribution and stats of everything
```{r}
#summary(cont_data)
```
 

- ID and EmployeeNumber, EmployeeCount,StandardHours should not be used for the model
  - StandardHours and Employee count only have one value, 80 for std hours and 1 for employee count
- People age ranges from 18-60 years old with a mean of 35
- Gender distribution
  - 59% Male and 41% female
  - More male and female
- Hourly rate and daily rate, MonthlyRate , standardHours seem to No correlation per graphs
- There seems to not be any relationship between Job involvement and job level
- Looking at those that have NumCompaniesWorked equal 0:
  - They have worked, it says minimum Years working at the company is 1, performance is mostly 3, they have a monthly income of mean $2,045 which means that these most likely are new employees who haven't completed a year.
  
```{r}
#are there duplicates?
fritolay[duplicated(fritolay)] # No Duplicates
```

```{r}
#remove col
fritolay=fritolay %>% select(-ID, -EmployeeNumber, -EmployeeCount, -NumCompaniesWorked, -StandardHours)

## Looking at those with NumCompnaiesWorked as 0
#fritolay %>% filter(NumCompaniesWorked==0)
#fritolay %>% filter(StockOptionLevel==0)
#fritolay %>% filter(TotalWorkingYears==0) #----->>>They seem to be new workers. They are 7,  18-19 years old, single with relationship between 3-4,age 18-19, companies worked 1 and it is their first company
#fritolay %>% filter(TrainingTimesLastYear==0)
#fritolay %>% filter(YearsAtCompany==0)
#fritolay %>% filter(YearsInCurrentRole==0)
#YearsSinceLastPromotion, yearswithcurrmanager, YearsInCurrentRole all 0
```

- Those with total working years as 0 seem to be new workers. They are 18-19 years old, single, with relationship between 3-4,age 18-19, companies worked 1, and their first company
- All zero's in the data make sense and most seem to portray new workers.


### Correlation
```{r, echo=FALSE,  fig.height = 15, fig.width = 15}

library("ggstatsplot")

ggstatsplot::ggcorrmat(
  data = cont_data,
  type = "parametric", # parametric for Pearson, nonparametric for Spearman's correlation
  colors = c("darkred", "white", "steelblue") # change default colors
)

```

**Normal trends that should appear in every company, are shown in the company**
  - Monthly Income vs Job Level
  - Performance Rate vs Percent Salary Hike level
  - Monthly Income vs Total Years working 
  - total working years and age 
  - Job level vs monthly income 
  - Monthly income vs total working years
  - Relationship vs Years With manager
  - Performance rating vs training
  - Jobrole vs Worklife balance
  - Performance Rating vs Relationship with manager
  - Job level and monthly income vs Total working years .78

- Research scientist are the happiest with their jobs. They have the highest Job Satisfaction, sales rep

# Interesting Facts
  - People that do not travel have tend to have the lowest overturn rate\
  - Research and development is the most prevalent department with 64% and it has the highest attriiton rate of 87%\
  -  Director jobs tend to have the lowest turn over rate. Only 2% quit their job
  - All people that studied Marketing are in sales, just like all people that studied human resources are in human resources. However thoe who had education in: Other, Technical, Science, Medicine work also work in other fields.
  - 17% of those that study medicine ended up going into Sales
  - 2.7% that did Technical deegree ended up in human resources which is surprising


# Linear regression to predict Monthly Income
```{r}
library(car)
library(lmtest)
library("car")
library(splitstackshape)


set.seed(20)#set seed
TrainObs = sample(seq(1,dim(fritolay)[1]),round(.75*dim(fritolay)[1]),replace = FALSE) #75% split
salaryTrain = fritolay[TrainObs,]
salaryTest = fritolay[-TrainObs,]
RMSE_store=c() #Sore all RMSE to compare
Model1_ = lm(MonthlyIncome ~ .-Department-Attrition, data = salaryTrain)
#Using with all the data

#vif(Model1_) #check collinearity
#The variance inflation is  high for JobRole. Due to the strong collinarity between Jobrole and another indepent variable, I need to remove it from model
#let's see the score though

### Important variables: YearsSinceLastPromotion, TotalWorkingYears,PerformanceRating, MonthlyRate,JobRoleLaboratory Technician,JobRoleManager ,JobRoleResearch Scientist, JobRoleResearch Director , JobLevel, BusinessTravelTravel_Rarely

Model1_ = lm(MonthlyIncome ~ .-Department-JobRole-Attrition, data = salaryTrain) #removing JobRole from the model as it has high collinarity
#vif(Model1_) #check collinearity again
#that is better so we can continue
#summary(Model1_)
#r sqr .90 not bad
Model1_Preds = predict(Model1_, newdata = salaryTest)#Predict
MSPE = mean((salaryTest$MonthlyIncome - Model1_Preds)^2)
RMSE <- function(error) { sqrt(mean(error^2)) }
RMSE(Model1_$residuals)#get RMSE
#1338.657

RMSE_store=c(RMSE_store,RMSE(Model1_$residuals))#store value

#Second try using most important values based on model one's p values
Model2_ = lm(MonthlyIncome~TotalWorkingYears+JobLevel+DistanceFromHome+YearsWithCurrManager,data=salaryTrain)
#vif(Model2_) #no collinearity between the dependent variables
#summary(Model2_)
#R .92
Model2_Preds = predict(Model2_, newdata = salaryTest) #predict
MSPE_2 = mean((salaryTest$MonthlyIncome - Model2_Preds)^2)
RMSE <- function(error) { sqrt(mean(error^2)) }
RMSE(Model2_$residuals) #get RMSE
#1351

#Try  model 2.5 using LOOCV
train(MonthlyIncome~TotalWorkingYears+JobLevel+DistanceFromHome+YearsWithCurrManager, method ="lm", data=fritolay, trControl = trainControl(method="LOOCV"))

#RMSE 1377

#Model 3 
Model3_ = lm(MonthlyIncome~TotalWorkingYears+I(TotalWorkingYears^2)+JobLevel+I(JobLevel^2)+DistanceFromHome+YearsWithCurrManager+I(YearsWithCurrManager^2), data=salaryTrain)
#vif(Model3_)#collinearity increased so it won't be a good model as we fails the assumptions of linear model
#summary(Model3_) #R .92
Model3_Preds = predict(Model3_, newdata = salaryTest) 
MSPE_3 = mean((salaryTest$MonthlyIncome - Model3_Preds)^2)
RMSE <- function(error) { sqrt(mean(error^2)) }
RMSE(Model3_$residuals)
#1259 <--better RMSE but too high correlation
#plot(Model3_)

RMSE_store=c(RMSE_store,RMSE(Model3_$residuals)) #save RMSE

#Model4 is better
train(MonthlyIncome~TotalWorkingYears+I(TotalWorkingYears^2)+JobLevel+I(JobLevel^2)+DistanceFromHome+I(DistanceFromHome^2)+YearsWithCurrManager+I(YearsWithCurrManager^2), method ="lm", data=fritolay, trControl = trainControl(method="LOOCV"))
#RMSE 1286
```

### Best model was Model was model 2 \
#### According to the p values produced, the model suggest that these are the most influencial variables when predicting monthly salary:
  - OverTimeYes                       < 2e-16 ***	
  - JobInvolvement                   1.20e-07 ***	
  - JobSatisfaction                  4.79e-06 ***	
  - MonthlyIncome                    0.000154 ***
  - MaritalStatusSingle              0.000412 ***
  - YearsSinceLastPromotion          0.004467 ** 
  
- Graph of Linear model results  
```{r}
new=salaryTest
new$Predictions=Model2_Preds
new %>% ggplot(aes(salaryTest$MonthlyIncome,Predictions)) + geom_point() + geom_smooth()+ ggtitle("Actual vs predictions")+ xlab("Actual")+
  ylab("Predicted")

RMSE_store=c(RMSE_store,RMSE(Model2_$residuals))#store value
```

##### Predict salary utilizing the best model

```{r}
#Now let's predict with our best model with not too much collinearity  Model 2
NoSalary=read.csv("CaseStudy2CompSet No Salary.csv")
Model2_Preds = predict(Model2_, newdata = NoSalary)
NoSalary$MonthlyIncome=Model2_Preds
NoSalary=NoSalary %>% select(ID,MonthlyIncome)
write.csv(NoSalary,file="NoSalaryResult.csv", row.names=FALSE) #creates new CSV with results

```



# NaibeBays to predict Attrition
```{r}
#Changing "Attrition" variable to a binary value
fritolay$AttritionBin<-ifelse(fritolay$Attrition=="Yes",1,0)
#get an idea on most important variables using lenear regression utilizing the categorical column of attrition
Model1_ = lm(AttritionBin ~ .-Attrition, data = fritolay)
#summary(Model1_)
#let's create a model with them and check
t2=lm(AttritionBin ~ OverTime+JobInvolvement+JobSatisfaction+MonthlyIncome+MaritalStatus+YearsSinceLastPromotion, data = fritolay)
#summary(t2)
#vif(t2) #not toomuch correlaiton! great! let's start with these variables

##Explore the density of Monthly income to standarize it
#ggplot(fritolay, aes(MonthlyIncome)) + geom_density(fill="blue")
#ggplot(fritolay, aes(log(MonthlyIncome))) + geom_density(fill="blue")
#ggplot(fritolay, aes(sqrt(MonthlyIncome))) + geom_density(fill="blue")

#Standarizing Monthly income with log as it seem to give the best distribution
fritolay$logMonthlyIncome<-log(fritolay$MonthlyIncome)

#Testing Accuracy of Naive Bayes model

library(e1071)
splitPerc = .70 #split 70%
trainIndices = sample(1:dim(fritolay)[1],round(splitPerc*dim(fritolay)[1]))
train = fritolay[trainIndices,]
test = fritolay[-trainIndices,]

#Test 1
model = naiveBayes(train,train$Attrition)
prediction = table(predict(model,test),test$Attrition)
CM = confusionMatrix(prediction, positive = "Yes")
CM

```
- Overfitted the data 

```{r}
#test 2
model2=naiveBayes(Attrition ~ OverTime+JobInvolvement+JobSatisfaction+MonthlyIncome+MaritalStatus+YearsSinceLastPromotion+logMonthlyIncome+EnvironmentSatisfaction+DistanceFromHome+WorkLifeBalance+JobRole+RelationshipSatisfaction, data=fritolay)
prediction2 = table(predict(model2,test),test$Attrition)
CM = confusionMatrix(prediction2, positive = "Yes")
CM

```

##### Predict on unlabled data

```{r}
#Predict on unlabeled data
NoAttrition=read.csv("CaseStudy2CompSet No Attrition.csv")
Attrition_Preds = predict(model2,NoAttrition)
NoAttrition$Attrition=Attrition_Preds
NoAttrition=NoAttrition %>% select(ID,Attrition)

#Create CSV
write.csv(NoAttrition,file="NoAttritionResult.csv", row.names=FALSE)
#read.csv("NoAttritionResult.csv")


#Verifying top 3 factors for attrition
#CM = confusionMatrix(prediction, positive = "Yes")
#m1 <- glm(AttritionBin~JobSatisfaction+OverTime+JobInvolvement, #family=binomial, data=fritolay)
#summary(m1)
```


### Graphs used to gather some of the above analysis and to understand data

```{r}
library(corrplot)
#pred <- Model1_ %>%predict_classes(salaryTest) %>% factor(0:8)

#help(predict_class)
 
#res_tab <- table(Pred = pred, Act = test_labels)
#res_prop <- prop.table(res_tab,2)
 
#author_key <- tibble(author = nt_frame$author, code = nt_frame$author_factor) %>%
 # unique %>%
#  arrange(code)
 
#colnames(res_prop) <- author_key$author
#rownames(res_prop) <- author_key$author
#corrplot(res_prop,is.corr = FALSE,
        # method = "circle", addCoef.col = "lightblue", number.cex = 0.7)

```

- Checking what constitute each department segment
```{r}
unique(fritolay %>% filter(Department=="Sales") %>% select(JobRole))
unique(fritolay %>% filter(Department=="Research & Development") %>% select(JobRole))
unique(fritolay %>% filter(Department=="Human Resources") %>% select(JobRole))
# Manager Appears on both Sales and Research Development so lets update the name
fritolay[fritolay=="Manager" & fritolay$Department=="Sales"]="Sales Manager"
fritolay[fritolay=="Manager" & fritolay$Department=="Human Resources"]="HR Manager"

```

- Education vs department insights
```{r}
library(viridis)
library("ggsci")

# Discrete color
fritolay %>% ggplot(aes (y=EducationField, fill=Department)) + geom_bar() + 
  ylab("Education Field") + 
  ggtitle("Educational field vs Department") 



fritolay %>% group_by(EducationField) %>% summarize(Total_people=n(), Sales=paste((0+(round(((sum(Department=="Sales")/n())*100),1))),"%",sep=""), Research=paste((0+(round(((sum(Department=="Research & Development")/n())*100),1))),"%", sep=""), HR=paste((0+(round(((sum(Department=="Human Resources")/n())*100),1))),"%", sep=""))

```

- All people that studied Marketing are in sales, just like all people that studied human resources are in human resources. However thoe who had education in: Other, Technical, Science, Medicine work also work in other fields.
- 17% of those that study medicine ended up going into Sales
- 2.7% that did Technical deegree ended up in human resources which is surprising

- Rest of graphs to understand data

```{r, echo=FALSE}

plot1=fritolay %>% ggplot(aes(x=DailyRate, fill="blue")) + 
  geom_histogram()+ggtitle("Daily Rate")  + theme(legend.position = "none") 
plot2=fritolay %>% ggplot(aes(x=MonthlyRate,fill="blue")) + 
  geom_histogram()+ggtitle("Monthly Rate") + theme(legend.position = "none") 
plot3=fritolay %>% ggplot(aes(x=HourlyRate,fill="blue")) + 
  geom_histogram()+ggtitle("Hourly Rate")  + theme(legend.position = "none") 

grid.arrange(plot3, plot1, plot2, ncol=3) 

plot1=fritolay %>% ggplot(aes(x=TrainingTimesLastYear, fill="blue")) + 
  geom_histogram()+ggtitle("TrainingTimesLastYear")  + theme(legend.position = "none") 
plot2=fritolay %>% ggplot(aes(x=YearsSinceLastPromotion, fill="blue")) + 
  geom_histogram()+ggtitle("YearsSinceLastPromotion")  + theme(legend.position = "none") 
plot3=fritolay %>% ggplot(aes(x=YearsWithCurrManager, fill="blue")) + 
  geom_histogram()+ggtitle("YearsWithCurrManager")  + theme(legend.position = "none") 

grid.arrange(plot3, plot1, plot2, ncol=3) 
```

```{r, echo=FALSE, out.width="180%", out.height="250%"}
a=gghistogram(fritolay, x = "YearsSinceLastPromotion", 
  add = "mean", rug = TRUE,
  fill = "Attrition", palette = c("darkturquoise", "ghostwhite")
  )
b=gghistogram(fritolay, x = "YearsWithCurrManager", 
  add = "mean", rug = TRUE,
  fill = "Attrition", palette = c("darkturquoise", "ghostwhite")
  )
c=gghistogram(fritolay, x = "YearsSinceLastPromotion", 
  add = "mean", rug = TRUE,
  fill = "Attrition", palette = c("darkturquoise", "ghostwhite")
  )
d=gghistogram(fritolay, x = "YearsInCurrentRole", 
  add = "mean", rug = TRUE,
  fill = "Attrition", palette = c("darkturquoise", "ghostwhite")
  )
e=gghistogram(fritolay, x = "EnvironmentSatisfaction", 
  add = "mean", rug = TRUE,
  fill = "Attrition", palette = c("darkturquoise", "ghostwhite")
  )
f=gghistogram(fritolay, x = "JobSatisfaction", 
  add = "mean",
  fill = "Attrition", palette = c("darkturquoise", "ghostwhite")
  )
grid.arrange(a,b,c,d,e,f, ncol=3)
```

- There also seems to not be any patterns between Rate columns and Attrition nor between Daily, montly and hourly rate.



```{r, echo=FALSE, out.width="180%", out.height="150%"}

p1=ggplot(fritolay, aes(x= as.factor(MaritalStatus),  group=Attrition)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.3) +
    labs(y = "Percent") +
    facet_grid(~Attrition)+ scale_y_continuous(labels = scales::percent) + xlab("MaritalStatus")+
  theme(legend.position = "none") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 3)) 

p2=ggplot(fritolay, aes(x= as.factor(Gender),  group=Attrition)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(..prop..), 
                   y= ..prop.. ), stat= "count", vjust = -.3) + 
    labs(y = "Percent") +
    facet_grid(~Attrition)+ scale_y_continuous(labels = scales::percent) +
    xlab("Gender") + theme(legend.position = "none") 
p3=ggplot(fritolay, aes(x= as.factor(BusinessTravel),  group=Attrition)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.3) +
    labs(y = "Percent") +
    facet_grid(~Attrition)+
    scale_y_continuous(labels = scales::percent) + xlab("BusinessTravel")+
  theme(legend.position = "none") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 3)) +
  theme(axis.text.x=element_blank())
p4=ggplot(fritolay, aes(x= as.factor(Department),  group=Attrition)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.1) +
    labs(y = "Percent") +
    facet_grid(~Attrition)+scale_y_continuous(labels = scales::percent) +
    xlab("Department") +
  theme(legend.position = "none") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 3)) 
p5=ggplot(fritolay, aes(y= as.factor(EducationField),  group=Attrition)) + 
    geom_bar(aes(x = ..prop.., fill = factor(..y..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(round(..prop..,2)),
                   x= ..prop.. ), stat= "count", vjust = -.3) +
    labs(x = "Percent") +
    facet_grid(~Attrition)+scale_x_continuous(labels = scales::percent) +
    ylab("EducationField")+
  theme(legend.position = "none") 
p6=ggplot(fritolay, aes(y= as.factor(JobRole),  group=Attrition)) + 
    geom_bar(aes(x = ..prop.., fill = factor(..y..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(round(..prop..,2)),
                   x= ..prop.. ), stat= "count", vjust = -.3) +
    labs(x = "Percent") +
    facet_grid(~Attrition)+scale_x_continuous(labels = scales::percent) +
    ylab("JobRole")+
  theme(legend.position = "none") 
p7=ggplot(fritolay, aes(x= as.factor(OverTime),  group=Attrition)) + 
    geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
    geom_text(size=3,aes( label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.3) +
    labs(y = "Percent") +
    facet_grid(~Attrition)+scale_y_continuous(labels = scales::percent) +
    xlab("OverTime")+
  theme(legend.position = "none") 
#grid.arrange(p2,p7,p1,p4,p3, ncol=2)
grid.arrange(p5,p6, ncol=1)
p1
```
```{r}
#new_day==cont_data %>% select("DailyRate")
#does not equals daily Rate!
ggplot(fritolay, aes(x=Age, fill=as.factor(JobSatisfaction))) + geom_bar() 
ggplot(fritolay, aes(y=YearsSinceLastPromotion,color=as.factor(JobSatisfaction))) + geom_boxplot() 

ggplot(fritolay, aes(x=YearsSinceLastPromotion, fill=as.factor(JobSatisfaction))) + geom_bar() 
cor.test(x=fritolay$JobSatisfaction, y=fritolay$YearsSinceLastPromotion,method = "spearman")
#no correclation btw years an las promotion

ggplot(fritolay, aes(x=YearsSinceLastPromotion, y=JobSatisfaction, color=as.factor(JobSatisfaction))) + geom_col() 
#environment satisfaction vs jobsatisfaction

ggplot(fritolay, aes(x=EnvironmentSatisfaction, y=JobSatisfaction, color=as.factor(JobSatisfaction))) + geom_col()
cor.test(x=fritolay$EnvironmentSatisfaction, y=fritolay$JobSatisfaction,method = "spearman")
#no correlation

ggplot(fritolay, aes(x=EnvironmentSatisfaction, fill=as.factor(JobSatisfaction))) + geom_bar()

ggplot(fritolay, aes(y=EnvironmentSatisfaction, fill=as.factor(JobSatisfaction))) + geom_boxplot()

ggplot(fritolay, aes(y=(JobRole), fill=as.factor(JobSatisfaction))) + geom_bar()

ggplot(fritolay, aes(y=(WorkLifeBalance), fill=as.factor(Attrition))) + geom_bar()
#no coee


l=fritolay  %>% group_by(JobRole) %>% summarize(one=round(((sum(JobSatisfaction==1)/n())*100),1),
                   two=round(((sum(JobSatisfaction==2)/n())*100),1),
                   three=round(((sum(JobSatisfaction==3)/n())*100),1),
                   four=round(((sum(JobSatisfaction==4)/n())*100),1)) 


ggplot(fritolay, aes(y= as.factor(JobRole),  group=JobSatisfaction)) + 
    geom_bar(aes(x = ..prop.., fill = factor(..y..)), stat="count") +
    geom_text(size=2,aes( label = scales::percent(..prop..),
                   x= ..prop.. ), stat= "count", vjust = -.3) +
    labs(y = "Percent") +
    facet_grid(~JobSatisfaction)+ scale_x_continuous(labels = scales::percent) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 3)) 
#there seems to not be a relationship between environment and job satisfaction

#monthly rate + job_satisfaction
ggplot(fritolay, aes(y=MonthlyRate, color=as.factor(JobSatisfaction))) + geom_boxplot() 
#the satisfaction for 3 and 4 have a higher mean monthly rate than that of 1 and 2

f=fritolay %>% group_by(EnvironmentSatisfaction) %>% summarize(one=round(((sum(JobSatisfaction==1)/n())*100),1),
                   two=round(((sum(JobSatisfaction==2)/n())*100),1),
                   three=round(((sum(JobSatisfaction==3)/n())*100),1),
                   four=round(((sum(JobSatisfaction==4)/n())*100),1)) 


```
